A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-03-18 graph_syntax.tex
tweak
update and merger

@@

Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. The family $\{A_i | i \in I\}$ is an indexed family of sets with index $i$ and indexing set $I$, and we can say $\mathcal A = \{A_i | i \in I\}$\cite{smith2006}. Then the union over this family of sets is
\begin{equation}
\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x | x \in A \txt{ for } A \in \mathcal A\}.
\end{equation}
Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$.

@@

Only the variable, model nodes are allowed to be used in an MCG. In a FPF expression 
nodes are also required. All four nodes are required for any SPF. In addition to 
the three node types, there are two edge types:

@@

\subsection{Fundamental Problem Formulation Graph}
We now define the fundamental problem formulation graph, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
\begin{enumerate}
\item[(1)] $\displaystyle{I_F \subset I}$ such that the following conditions hold
\item[(2)] $\displaystyle{ \forall i \in I, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} | \txt{deg}^+(v) > 0\} }$
\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\end{enumerate}
The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). Requirement (2) stipulates that the analysis blocks in $F$ must each have at least one local output that is beging used. Requirement (3) stipulates that only the global inputs that are being used should be included. Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.

@@

\begin{enumerate}
\item The first step is to detect holes and disconnect the free edges following them. These free edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs. The set of variable nodes which are holes is created as
\begin{equation}
H = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
\end{equation}
which is the set of variable nodes with few incoming edges than are allowed by the lower indegree limit.
Then the updated set of edges is created by removing the first set of free edges following the hole:
\begin{equation}
C_{F,1} = C_{F,0} \setminus \{(x,y) \in E | x=E^3(v) \txt{ for } v \in H\},
\end{equation}
%deg-v is now recaculated for the new C. 
and this step is demonstrated by Fig.~\ref{f:holes}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_hole}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a hole.}
\label{f:holes}
\end{figure}
\item The second step is to detect collisions and to then disconnect enough free edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
\begin{equation}
S_\txt{nodes} = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
\end{equation}
For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
\begin{equation}
S_\txt{edges} = \big \{ \{(x,y) \in E\} \big | y \in S_\txt{nodes} \big \}
\end{equation}
Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$. Then we may construct sets of edges as
\begin{equation}
B_j = \big \{e_k \in S_{\txt{edges},j} \ | \ k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
\end{equation}
which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$. The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
\begin{equation}
C_{F,2} = \{ e \in C_{F,1} \ | \ e \in B_j \txt{ for some } j \in J\}.
\end{equation}
\item The third and final step is to remove any unused nodes and edges. The set of unused variable nodes representing local outputs is
\begin{equation}
V_\txt{empty} = \{v \in V_M \ | \ t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) = 0\},
\end{equation}
and so then the indexing set for only the analysis blocks with at least one used local output may be constructed as
\begin{equation}
I_3 =  \{i \in I \ | \ \nexists v \in V_{A_i} \txt{ such that }  v \in V_\txt{empty} \}.
\end{equation}
Finally, the set of nodes and edges describing a fundamental problem formulation is given by
\begin{equation}
V_{F,3} = V_M \setminus V_\txt{empty}
\end{equation}
\begin{equation}
C_{F,3} = \{ (x,y) \in C_{F,2} \ | \ x \in V_{A_i} \txt{ or } y \in V_{A_i} \txt{ for some } i \in I_3 \}
\end{equation}
\begin{equation}
F_3 = (V_{F,3},C_{F,3})
\end{equation}

>>>>

\newcommand{\st}{ \ | \ }

@@

Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. The family $\{A_i \st i \in I\}$ is an indexed family of sets with index $i$ and indexing set $I$, and we can say $\mathcal A = \{A_i \st i \in I\}$\cite{smith2006}. Then the union over this family of sets is
\begin{equation}
\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x \st x \in A \txt{ for } A \in \mathcal A\}.
\end{equation}
Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$.

@@

Only the variable, model, and expression nodes are allowed to be used in an MCG or FPF. 
All four nodes are required for any SPF. In addition to the three node types, there are two edge types:

@@

\subsection{Fundamental Problem Formulation Graph}
We now define the fundamental problem formulation graph, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
\begin{enumerate}
\item[(1)] $\displaystyle{I_F \subset I}$ such that the following conditions hold
\item[(2)] $\displaystyle{ \forall i \in I, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\end{enumerate}
The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). Requirement (2) stipulates that the analysis blocks in $F$ must each have at least one local output that is beging used. Requirement (3) stipulates that only the global inputs that are being used should be included. Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.

@@

\begin{enumerate}
\item The first step is to detect holes and disconnect the free edges following them. These free edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs. The set of variable nodes which are holes is created as
\begin{equation}
H = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
\end{equation}
which is the set of variable nodes with few incoming edges than are allowed by the lower indegree limit.
Then the updated set of edges is created by removing the first set of free edges following the hole:
\begin{equation}
C_{F,1} = C_{F,0} \setminus \{(x,y) \in E \st x=E^3(v) \txt{ for } v \in H\},
\end{equation}
%deg-v is now recaculated for the new C. 
and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_hole}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a hole.}
\label{f:hole}
\end{figure}

\item The second step is to detect collisions and to then disconnect enough free edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
\begin{equation}
S_\txt{nodes} = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
\end{equation}
For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
\begin{equation}
S_\txt{edges} = \big \{ \{(x,y) \in E\} \st y \in S_\txt{nodes} \big \}
\end{equation}
Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. An examle collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$. Then we may construct sets of edges as
\begin{equation}
B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
\end{equation}
which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$. The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
\begin{equation}
C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\}.
\end{equation}
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_collision}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a collision.}
\label{f:collision}
\end{figure}

\item The third and final step is to remove any unused nodes and edges. The set of unused variable nodes representing local outputs is
\begin{equation}
V_\txt{empty} = \{v \in V_M \st t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) = 0\},
\end{equation}
and so then the indexing set for only the analysis blocks with at least one used local output may be constructed as
\begin{equation}
I_F =  \{i \in I \ | \ \nexists v \in V_{A_i} \txt{ such that }  v \in V_\txt{empty} \}.
\end{equation}
The set of used global inputs is constructed as
\begin{equation}
V_{\txt{in},F} = \{v \in V_\txt{in} \st |\{(x,y) \in E(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1\},
\end{equation}
which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be two the analysis blocks indexed by $I_F$.

Finally, the set of nodes and edges describing a fundamental problem formulation is given by
\begin{equation}
V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
\end{equation}
\begin{equation}
C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
\end{equation}
\begin{equation}
F = (V_F,C_F).
\end{equation}
